\normalsize{Biometry of the invasive lionfish (\textit{Pterois volitans}) in the Central Mexican Caribbean, and a review of allometric growth parameters across the invasion range}

### 2017-08-04 Manuscript.Rmd
Updated manuscript for submission

---
title: \normalsize{Biometry of the invasive lionfish (\textit{Pterois volitans}) in
  Playa del Carmen, Mexico and a review of length - weight parameters across the invasion
  range}
csl: apa.csl
fontsize: 12pt
output:
  pdf_document:
    fig_caption: yes
header-includes:
- \usepackage{setspace}
- \doublespacing
- \usepackage{lineno}
- \linenumbers
bibliography:
- references.bib
- packages.bib
---

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: { 
      equationNumbers: { 
            autoNumber: "all",
            formatNumber: function (n) {return '9.'+n}
      } 
  }
});
</script>

@@

@@

Juan Carlos Villaseñor-Derbez^1^\

^1^ Bren School of Environmental Sciences and Management, University of California Santa Barbara\

**Full postal address:**
4312 Bren Hall, Santa Barbara, CA, 93106\

**Running title:** Length - weight parameters for *Pterois volitans* \

*Corresponding author information:*

Juan Carlos Villaseñor-Derbez   
4312 Bren Hall  
Santa Barbara, CA   
93106   
jvillasenor@bren.ucsb.edu   
+1 (207) 205 8435

@@

**Title of the contribution:** Biometry of the invasive lionfish (*Pterois volitans*) in Playa del Carmen, Mexico and a review of length - weight parameters across the invasion range\

**Abstract**

300 words\

**Key words**

lionfish, biometry, length-weight relationship, Mexico\

**Resumen**

300 palabras\

**Palabras clave**

pez león, biometría, relación lopngitud-peso, México\

@@

At least 84% of the marine eco-regions have reported the presence of an invasive species [@molnar_2008]. The presence of invasive alien species represents a major threat to local biodiversity and the economic activities that depend on it, like tourism or fisheries [@bax_2003]. By 2005, the economic cost of invasive species to the United States was  estimated at \$120 billion per year and nearly 42% of species that have been included in the Endangered or Threatened species list have been labeled as such due to presence of invasive species [@pimentel_2005]. Invasive species may threaten native species through competition [@davis_2003] or predation. This highlightes the importance of understanding, managing, and preventing ecological invasions.

Lionfish (*Pterois volitans - miles* complex) are an invasive species in the North-Western Atlantic and the Caribbean likely introduced through libreation of aquarium-kept organisms [@betancurr_2011]. They are the first marine vertebrates to establish in North Atlantic [@schofield_2009: @schofield_2010] and Caribbean coasts. Lionfish have been recorded in coral reefs [@aguilarperera_2010], but also in other habitats like estuaries [@jud_2011], mangroves [@barbour_2010] and hard-bottom reefs [@muoz_2011]. Their presence in these waters has been labeled as a major marine invasion due to its threat to local biodiversity, the veolocity of its invasion, and its difficulty of management [@hixon_2016].

As ways to reduce the abundance of this invasive species local authorities have promoted removal programs incentivized consumption of this invasive species [@chin_2016]. Lionfish removals have shown to significantly reduce -but not quite eliminate- lionfish abundances at local scales [@sandel_2015, @chin_2016, @deleon_2013]. While complete erradication through fishing effort is unlikely to be a solution due to the rapid recovery rates exhibited by the species [@barbour_2011], further incentivizing its consumption might create a market demand big enough to promote a fishery [@chin_2016].

A significant amount of research has been done on lionfish feeding ecology in North Carolina [@muoz_2011], the Bahamas [@morris_2009; @cote_2013], Northern Gulf of Mexico [@dahl_2014], Mexican Caribbean [@valdezmoreno_2012; @villaseorderbez_2014], and Costa Rica [@sandel_2015]. It has been identified that lionfish can reduce recruitment [@albins_2008] and population sizes [@green_2012] of native reef-fish species, and further the endangerement of critically endangered reef fish [@rocha_2015]. This information has been useful in predicting the trophic impacts of lionfish [@ariasgonzalez_2011]. Significant efforts have also been made to understand the possible impacts of the invasion by keeping track of the invasion range through time [@schofield_2009; @schofield_2010] and predicting invasion ranges under climate change scenarios (@grieve_2016).

Biomass is a standard indicator used to measure the state or size of a population. Biomass has been used in lionfish research to convert individual fish length observations into fish weight. In order to do this, researchers must find growth parameters thad describe the growth of their species for a particular region or location. Having precision of these parameters is important, because small changes in any of them can result in important over- or underestimation of biomass levels. This becomes particularly important when research involves, for example, identifying the total biomass available for harvest by fishers -as invasion management programs- or the efficacy of lionfish removals.

The objectives of this study include providing site-specific weight-at-length parameters for the invasive lionfish in the Mexican Caribbean region. At the same time, we highlight the importance of using site-specific parameters by estimating biomass with parameters from other regions across the invasion range and comparing them to observed biomasses. Finally, we summarize other weight-at-length studies presented, making them readily accessible for future research.

@@

The present study took place off the coasts of Playa del Carmen, in the coasts of the Mexican Caribbean. The region represents the northernmost section of the Mesoamerican reef, which extends from coast of Cancun south to Honduras (**reference**). Coral reefs and mangroves are important habitats distributed along the cost, and represent important sources of income in terms of extractive (*e.g.* recreational fishing) and non-extractive (*e.g.* SCUBA diving) activities related to tourism (**reference**), the main source of income to the local economy.

Coral reefs in the region are characterized by X, Y, Z. Descripcion general de los arrecifes.

@@

The present study used samples obtained by @villaseorderbez_2014 for stomach content analysis between May and August of 2010. A total of 33 SCUBA immersions were performed in `r length(unique(fish$Location))` sampling sites along 14 Km of coast between Puerto Aventuras and Akumal (Fig. 1, Table 1). Sampling locations included wall and carpet reefs at depths between `r min(fish$Depth_m)` m and `r max(fish$Depth_m)` m.  All observed organisms (n = `r dim(fish)[1]`) were collected using hand nets -to avoid weight loss due to bleeding- and numbered collection bottles. Information on depth and other comments were recorded in an underwater slide. Depth was recorded by dive gauges held by divers as safety procedures during the collections. Samples were frozen within 30 minutes of completing the dive and stored for posterior analysis in the lab. For every organism, Total Length (TL; mm) and Total Weight (TW; gr) were recorded in the lab.

@@

location_table <- fish %>% 
  group_by(Location) %>% 
  summarize(Latitude = formatC(mean(Latitude), digits = 3, format = "f"),
            Longitude = formatC(mean(Longitude), digits = 3, format = "f"),
            minD = formatC(min(Depth_m), digits = 1, format = "f"),
            maxD = formatC(max(Depth_m), digits = 1, format = "f"),
            meanD = formatC(mean(Depth_m), digits = 1, format = "f"),
            n = n()) %>% 
  rbind(total) %>% 
  knitr::kable(row.names = F, caption = "Coordinates, minimum, maximum and mean depth (m), and number of samples for each location. n = sample size.", col.names = c("Location", "Lat.", "Long.", "Min. Depth", "Max. Depth", "Mean Depth", "n"))

@@

The weight at length relationship between the observed variables was calculated with the allometric growth function:

@@

\begin{equation}log_{10}(TW) = blog_{10}TL + log{10}(a)\end{equation}

@@

Where $Y = log_{10}(TW)$, $X = log_{10}(TL)$, $m = b$, and $c = log_{10}(a)$. The coefficients ($c$ and $b$) were estimated with  an Ordinary Least Square Regression and heteroskedastic-robust standard errors. Both coefficients were tested agains the null hypothesis of no change (*i.e.* $h_0: c = 0$ and $h_0: b = 0$). Additionally, the allometric parameter was tested against the null hypothesis of isometric growth ($H_0: b = 3$). Coefficient testing was made with a two-tailed Student's t-test. The significance of the regression was corroborated with an F-test.

Other growth parameters were extracted from published research from North Carolina [@barbour_2011], Northern [@fogg_2013] and Southern [@aguilarperera_2016] Gulf of Mexico, the Southern Mexican Caribbean [@sabidoitza_2016], Little Cayman [@edwards_2014], Jamaica [@chin_2016], Bonaire [@deleon_2013] and Costa Rica [@sandel_2015]. Additionally, parameters from Fishbase [@froese_website_2016] were also included in this study. When available, information on sampling methods, gender differentiation, location, and depth ranges of each study was retreived. Whenever gender was not specified, it was assumed that the results were presented for both genders pooled together. During this review process, it was noticed that some papers indistinctly use $a$ to report either the ponderal index in **eq. 1** or the y-intercept ($c$) in **eq. 2**, which might sometimes be overlooked. Furthermore, some studies report their parameters as mm-to-gr conversions, but a rapid evaluation of such parameters indicates that they were estimated as cm-to-gr conversions. Here, both parameters ($a$ and $c$) are reported for the present findings and, when ever required, values from other studies are converted for consistency. All coefficients are reported as mm to gr conversions.

Mean observed biomass was calculated as the average Total Weight of all observed organisms. Expected biomass for each organism was calculated with the growth parameteres estimated in this study, as well as those retreived from additional literature. Then, mean expected biomass was calculated for each study. All expected biomass values were divided by the observed biomass to obtain a ratio that allowed rapid identification of over- or underestimation with respect to the observed biomass in this study. Therefore, values lower than 1 indicate that the parameters used underestimate the biomass. On the other side, values larger than 1 indicate that the parameteres used overestimate biomass. Median and mean expected to observed biomass ratios were calculated across all sites for each study.

All hypothesis testing was performed with a confidence level of $\alpha = 0.01$ and performed in R version 3.4.0 [@rcore_2017]. Data wrangling was performed with the tidyverse package [@tidyverse_2017]. Maps were created with a mix of functions from the sp [@sp_2017], rgdal [@rgdal_2017], tmap [@tmap_2017], and tmaptools [@tmaptools_2017] packages. Correction for heteroskedastic-robust standard errors was done with the sandwich [@sandwich_2014] and lmtest [@lmtest_2002] packages. Models were manipulated with with the broom package [@broom_2017]. Stargazer [@stargazer_2015] was used to produce the tables, and refmanager [@RefManager_2014] was used to keep track of citations. The manuscript was written in rmarkdown [@rmarkdown_2017] and processed with the knitr package [@knitr_2017].

@@

A summary of Total Lengths and Total Weights of the the sampled organisms (n = `r dim(fish)[1]`) is presented in Table 2. The smallest organism (TL = `r formatC(min(fish$Length), digits = 2, format = "f")` mm) was also the lightest organism (TW = `r formatC(min(fish$Weight), digits = 2, format = "f")` gr). However, the largest organism (TL = `r formatC(max(fish$Length), digits = 2, format = "f")` mm) was not the heaviest (TW = `r formatC(fish$Weight[fish$Length == max(fish$Length)], digits = 2, format = "f")` gr), and the heaviest organism (TW = `r formatC(max(fish$Weight), digits = 2, format = "f")` gr) was `r formatC(fish$Length[fish$Weight == max(fish$Weight)], digits = 2, format = "f")` mm in total length. Kernell density plots (Fig. 2) show the distribution for TL and TW. Both measures were positively skewed, with skewness of `r formatC(skewness(fish$Length), digits = 2, format = "f")` for TL and `r formatC(skewness(fish$Weight), digits = 2, format = "f")` for TW.

```{r summary table, include = F}
summary_table <- fish %>% 
  select(Length, Weight) %>% 
  as.data.frame() %>% 
  stargazer(header = F, type = "latex", digits = 2, title = "Summary statistics of Length and Weight of sampled organisms.")
```

@@

wald_results <- waldtest(model, vcov = vcovHC(model, type = "HC1"))
f_string <- c("F Statistic", paste0(formatC(wald_results$F[2], digits = 2, format = "f"), "*** (df = 1; ", dim(fish)[1]-2,")"))

@@

The model adjusted to eq. 3 estimated the coefficient values at $m = `r coef(model)[[2]]`$ and $c = `r coef(model)[[1]]`$. The intercept and slope were significantly different from zero ($t(107) = `r formatC(coef(model)[[1]]/robust_se[[1]], digits = 2, format = "f")`; p<0.01$ and $t(107) = `r formatC(coef(model)[[2]]/robust_se[[2]], digits = 2, format = "f")`; p<0.01$, respectively), rejecting the null hypothesis of no change. Additionally, the allometric factor was significantly different from zero ($t(107) = `r formatC((coef(model)[[2]]-3)/robust_se[[2]], digits = 2, format = "f")`; p<0.01$). More information on model fit and confidence intervals for the estimated coefficients are presented in Table 3. The relationship between Total Length and Total Weight is presented in Figure 3.

```{r regression table, include = F}
regression_table <- stargazer(model,
          header = F,
          se = list(robust_se),
          type = "latex",
          t.auto = T,
          p.auto = T,
          intercept.bottom = F,
          single.row = T,
          report = "vcs*",
          omit.stat = c("rsq", "f"),
          add.lines = list(f_string),
          ci = T)
```

@@

```{r, include = F}
review <- read_csv("biometric_review.csv", col_types = cols())

review_table <- review %>% 
  select(Reference, n = N, a, c, b, r2, Gender, minD, maxD, Spear) %>% 
  knitr::kable(row.names = F, caption = "Allometric growth parameters for eight published papers, Fishbase (Froese and Pauly 2016), and this study. All parameters have been adjusted to convert from millimiters to grams. n = Sample size, a = scaling parameter for eq. 1, c = y-intercept for eq. 3, b = exponent or slope for eq. 1 or eq. 2, respectively.", digits =  c(1, 0, 6, 4, 4, 4, 0, 1, 1))

review_table
```

From the eight peer-reviewd studies including information on growth parameters for *P. volitans* and FishBase [@froese_website_2016], `r dim(review)[1]-1` parameters were identified. Two studies [@aguilarperera_2016; @fogg_2013] reported gender-level and pooled parameters, while the rest of the studies always presented pooled results. The smallest coefficient of determination was presented by @chin_2016 with $R^2 = `r min(review[,12], na.rm = T)`$, while @sabidoitza_2016 reported the highest value at $R^2 = `r min(review[,12], na.rm = T)`$. These studies presented information for organisms obtained at depths between `r min(review$minD, na.rm = T)` and `r max(review$maxD, na.rm = T)` meters. Two studies [@aguilarperera_2016; @chin_2016] explicitly stated that their organisms were sampled with pole spears. Five studies [@sandel_2015; @barbour_2011; @fogg_2013; @edwards_2014; @sabidoitza_2016] mentioned that some of their organisms were obtained with pole spears (or other type of harpoon). A single study [@deleon_2013] did not specify how samples were obtained.

Parameters from models fit to males or females exclusively tend to have a higher steapness (*i.e.* higher allometric parameter), with mean $\pm$ standard deviation values of $b = `r formatC(mean(review$b[review$Gender == "M"]), digits = 2, format = "f")` \pm `r formatC(sd(review$b[review$Gender == "M"]), digits = 2, format = "f")`$ and $b = `r formatC(mean(review$b[review$Gender == "F"]), digits = 2, format = "f")` \pm `r formatC(sd(review$b[review$Gender == "F"]), digits = 2, format = "f")`$, compared to parameters from models for pooled genders with a mean $\pm$ standard deviation value of $b = `r formatC(mean(review$b[review$Gender == "Both"]), digits = 2, format = "f")` \pm `r formatC(sd(review$b[review$Gender == "Both"]), digits = 2, format = "f")`$. In the case of the ponderal index ($a$) and its $log_{10}$ transformed parameter ($c$), values were higher for parameters for pooled genders. The trends in $a$, $c$, and $b$ indicate that some models will tend to estimate Total Weights differently. Figure 4 shows the predicted weights for organisms within the size range of these study using the 14 parameters previously described.

@@

# Discussion and Conclusions

Wider range sizes

Model fit and this model

No spear doesnt seem to be a determinant factor in weight, but spear doesnt allow catching small organisms. Not using spear poles allows us to have a full sample of fish with a wider range of sizes and weights, ideal for visual census. Also, there is no loss in body mass due to bleeding.

Gender-level overestimates and is steaper because it lacks small organisms to pull the curve

>>>>

---
title: \normalsize{Biometry of the invasive lionfish (\textit{Pterois volitans}) in the Central Mexican Caribbean, and a review of allometric growth parameters across the invasion range}
fontsize: 12pt
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes:
- \usepackage{setspace}
- \doublespacing
- \usepackage{lineno}
- \linenumbers
bibliography:
- references.bib
- packages.bib
csl: huct.csl
---
<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: { 
      equationNumbers: { 
            autoNumber: "all",
            formatNumber: function (n) {return '9.'+n}
      } 
  }
});
</script>

@@

options(knitr.table.format = "latex")

@@

Juan Carlos Villaseñor-Derbez^1^

^1^ Bren School of Environmental Sciences and Management, University of California Santa Barbara, Santa Barbara, California, U.S.

**Full postal address**   
Bren Hall   
University of California at Santa Barbara   
Santa Barbara   
CA, 93106

**Corresponding author information:**   
Juan Carlos Villaseñor-Derbez   
Bren Hall   
University of California at Santa Barbara   
Santa Barbara   
CA, 93106
jvillasenor@bren.ucsb.edu
+1 207 205 8435

**Running title**    
Allometric growth of Pterois volitans

@@

**Title of contribution**: Biometry of the invasive lionfish (*Pterois volitans*) in the Central Mexican Caribbean, and a review of allometric growth parameters across the invasion range

**Abstract**    
Lionfish (*Pterois volitans/miles*) are an invasive species in the North-Western Atlantic and the Caribbean, likely introduced through liberation of aquarium-kept organisms. Due to its threat to local biodiversity, the speed of their spread, and its difficulty of management, their presence in these waters has been labeled as a major marine invasion. In order to better manage the invasion, inform lionfish removal programs, predict their possible impacts, and estimate biomass available for harvest, we must be able to accurately estimate their total biomass. This study reports allometric growth parameters ($a = 3.2056\times 10^{-6}; b = 3.235$) for organisms sampled in the central portion of the Mexican Caribbean. It also identifies that, for a given length, organisms from the Caribbean tend to have lower weights than those from the Gulf of Mexicoand Atlantic, indicating that weight-at-length is spacialy variable. In addition to the new pair of parameters provided in this study, other 13 pairs of allometric growth parameters were used to calculate expected biomass from total length observations. Mean expected-to-observed biomass ratios by study ranged between 0.98 and 1.75, with overall values ranging from 0.44 to 3.51. The large differences between observed and expected biomass when using parameters from other locations highlights the importance of using site-specific parameters to estimate biomass from length observations. This becomes increasingly important when estimating biomass available for harvest or effectiveness of lionfish removal efforts. Additionally, guidelines on reporting allometric parameters are provided as a suggestion for future research are provided.

**Key words**   
lionfish, Pterois volitans, length-weight relationship, allometric growth, Mexico

**Resumen**   
Los peces león (*Pterois volitans/miles*) son una especiei nvasora en el Atlántico Noroeste y el Caribe, probablemente introducidos por medio de la liberación de organismos de ornato. Dada la amenaza que estos representan para la biodiversidad local, la velocidad de su invasión y dificultad de manejo, su presencia ha sido catalogada como una importante invasión marina. Para manejar mejor la invasión, proveer información a los programas de remoción, predecir los posibles impactos y estimar biomasa disponible para la extracción, debemos ser capaces de estimar su biomasa con precisión. Este estudio presenta parámetros de crecimiento alométrico ($a = 3.2056\times 10^{-6}; b = 3.235$) para organismos muestreados en la porción central del Caribe Mexicano. También identifica que para una misma longitud, los organismos del Caribe tienden a ser menos pesados que aquellos del Golfo de México y Atlántico, indicando variaciones espaciales en la relación peso-longitud. Además de los parámetros descritos en este estudio, se utilizaron otros 13 pares de parámetros de crecimiento alométrico para estimar biomasa a partir de observaciones de talla. La razón entre biomasa observada y biomasa esperada varió, en promedio por estudio, entre 0.98 y 1.75, con valores extremos de 0.44 y 3.51. Las grandes diferencias entre biomasa observada y esperada al usar parámetros de otras regiones resalta la importancia de utilizar parámetros específicos a la zona para estimar biomasa a partir de observaciones de talla. Adicionalmente, se proveen sugerencias para estandarizar la manera en que los parámetros de crecimiento alométrico se reportan.

**Palabras clave**
pez león, Pterois volitans, relación longitud-peso, crecimiento alométrico, México

@@

At least 84% of the marine eco-regions have reported the presence of an invasive species [@molnar_2008]. These represent a major threat to local biodiversity and the economic activities that depend on it, like tourism or fisheries [@bax_2003]. Invasive species may also threaten native species through competition [@davis_2003] or predation. By 2005, the economic cost of invasive species to the United States was  estimated at \$120 billion per year and nearly 42% of species that have been included in the Endangered or Threatened species list have been labeled as such due to presence of invasive species [@pimentel_2005]. This highlights the importance of understanding, managing, and preventing ecological invasions.

Lionfish (*Pterois volitans/miles* complex) are an invasive species in the North-Western Atlantic and the Caribbean, likely introduced through liberation of aquarium-kept organisms [@betancurr_2011]. They are the first marine vertebrates to establish in North Atlantic [@schofield_2009; @schofield_2010] and Caribbean coasts [@sabidoitza_2016]. Lionfish have been widely reported in coral reefs [@aguilarperera_2010], but also in other habitats such as estuaries [@jud_2011], mangroves [@barbour_2010], areas with hard-bottoms [@muoz_2011], and mesophotic reefs [@andradibrown_2017]. Due to its threat to local biodiversity, the speed of their spread, and its difficulty of management, their presence in these waters has been labeled as a major marine invasion  [@hixon_2016].

A significant amount of research has been done to describe lionfish feeding ecology in North Carolina [@muoz_2011], the Bahamas [@morris_2009; @cote_2013], Northern Gulf of Mexico [@dahl_2014], Mexican Caribbean [@valdezmoreno_2012; @villaseorderbez_2014], Belize [@hackerott_2017], and Costa Rica [@sandel_2015]. Their feeding behavior and high consumption rates can reduce recruitment [@albins_2008] and population sizes [@green_2012] of native reef-fish species, and further the endangerment of critically endangered reef fish [@rocha_2015]. (However, see @hackerott_2017 for a case where there was no evidence that lionfish affected the density, richness, or composition of prey fishes). Major efforts have also been made to understand the possible impacts of the invasion by keeping track of its range through time [@schofield_2009; @schofield_2010] and predicting invasion ranges under climate change scenarios [@grieve_2016]. By combining information from these disciplines, researchers have been able to predict the trophic impacts of lionfish [@ariasgonzalez_2011], which can then be translated into ecosystem-level and economic impacts.

Seeking to reduce lionfish densities, governments and non-profit organizations have promoted removal programs and incentivized its consumption [@chin_2016]. In some cases, these have shown to significantly reduce -but not quite eliminate- lionfish abundances at local scales [@sandel_2015, @chin_2016, @deleon_2013]. The rapid recovery rates exhibited by lionfish [@barbour_2011] and the persistent populations in mesophotic coral ecosystems [@andradibrown_2017] -which can contribute with recruitment to shallow-water populations- make of complete eradication through fishing effort an unlikely solution. However, further incentivizing its consumption might create a demand big enough to promote and sustain a stable fishery [@chin_2016], which can reduce local abundances and control -not eradicate- the invasion while providing alternative livelihoods.

The feasibility of lionfish removal programs has been extensively evaluated through field observations[@usseglio_2017; @sandel_2015, @chin_2016, @deleon_2013] and empirical modeling [@barbour_2011; @morris_2011; @johnston_2015]. The latter measure changes in biomass or density [@barbour_2011; @johnston_2015] in response to increased mortality (*i.e.* lionfish removal). In this case, biomass represents the sum of all fish's individual weight. Total Weight (TW) can be estimated from Total Length (TL) observations using the allometric growth equation ($TW = aTL^b$). Parameters $a$ and $b$ for this equation exist for North Carolina [@barbour_2011], Northern [@fogg_2013] and Southern Gulf of Mexico [@aguilarperera_2016], the Southern Mexican Caribbean [@sabidoitza_2016], Little Cayman [@edwards_2014], Jamaica [@chin_2016], Bonaire [@deleon_2013] and Costa Rica [@sandel_2015], but remain unavailable for the central Mexican Caribbean. The weight-at-length of a species can vary across regions as a response to biotic (*e.g.* local food availability) and abiotic (*e.g.* water temperature) conditions [@johnson_2016]. Thus, when using biomass-informed models or estimating biomass from length observations, it is important to use site-specific parameters to obtain an accurate estimate. This is especially important when research involves identifying the total biomass available for harvest by fishers [@chin_2016] or the efficacy of lionfish removals [@barbour_2011; @morris_2011; @johnston_2015].

Here, I provide the first allometric growth parameters for the invasive lionfish in the central Mexican Caribbean region. At the same time, I highlight the importance of using site-specific parameters by estimating biomass with parameters from other regions across the invasion range and comparing them to observed biomass. I also provide other 13 standardized parameters from eight studies through the invasion range, making them readily accessible for future research. Finally, I discuss the way in which allometric parameters are reported, and call for standardization to facilitate their use.

@@

The study took place off the coasts of Playa del Carmen, in the Mexican Caribbean (**Fig. 1**). The region represents the northernmost section of the Mesoamerican Barrier Reef System [@ruizzarate_2004]. Coral reefs and mangroves are locally important habitats that represent important sources of income in terms of extractive (*e.g.* recreational fishing) and non-extractive (*e.g.* SCUBA diving) activities related to tourism, the main source of income to the local economy [@murray_2007].

In the region, the reef profile has been described by @ariasgonzalez_1998, indicating that the reef lagoon extends about 500 m from the coast, until the reef crest is reached. The reef becomes deeper, leading to the reef front often found at 700 m from the coastline end extends for an additional 300 m. At approximately 1000 m away from shore and 30 - 40 m depth, the reef leads to a drop-off. Along a perpendicular profile to the coast, bands of reef are interrupted by sand patches at 8 - 12 m deep and 16-18 m deep. Along the coast, these reefs have been reported to be under significant anthropogenic pressure, likely causing a shift in structure and function [@bozec_2008].

@@

Organisms sampled by @villaseorderbez_2014 for stomach content analysis between May and August of 2010 were used in this work. A total of 33 SCUBA immersions were performed in `r length(unique(fish$Location))` sampling sites along the coast (Fig. 1, Table 1). Sampling locations included wall and carpet reefs at depths between `r min(fish$Depth_m)` m and `r max(fish$Depth_m)` m.  All observed organisms (n = `r dim(fish)[1]`) were collected using hand nets and numbered collection bottles. The use of hand nets prevents any weight loss due to bleeding(when collected with spears), and increases the range of sizes that can be sampled, providing a representative sample of the population. Information on depth and other comments were recorded in an underwater slide. Depth was recorded by dive gauges held by divers as safety procedures during the collections. Samples were frozen within 30 minutes of completing the dive and stored for posterior analysis in the lab. For every organism Total Length (TL; mm) and Total Weight (TW; gr) were recorded in the lab.

@@

location_table <- fish %>% 
  group_by(Location) %>% 
  summarize(Latitude = formatC(mean(Latitude), digits = 3, format = "f"),
            Longitude = formatC(mean(Longitude), digits = 3, format = "f"),
            minD = formatC(min(Depth_m), digits = 1, format = "f"),
            maxD = formatC(max(Depth_m), digits = 1, format = "f"),
            meanD = formatC(mean(Depth_m), digits = 1, format = "f"),
            n = n()) %>% 
  rbind(total) %>% 
  knitr::kable(row.names = F,
               col.names = c("Location", "Lat.", "Long.", "Min. Depth", "Max. Depth", "Mean Depth", "n"),
               booktabs = T)

@@

The weight at length relationship between the observed variables is described by the allometric growth function:

@@

\begin{equation}log_{10}(TW) = b\times log_{10}(TL) + log{10}(a)\end{equation}

@@

Where $Y = log_{10}(TW)$, $X = log_{10}(TL)$, $m = b$, and $c = log_{10}(a)$. Since $b == m$, we will only use $b$ throughout the paper for simplicity. The coefficients ($c$ and $b$) were estimated with  an Ordinary Least Square Regression and heteroskedastic-robust standard errors. Both coefficients were tested against the null hypothesis of no change (*i.e.* $H_0: c = 0$ and $H_0: b = 0$). Additionally, the allometric parameter was tested against the null hypothesis of isometric growth ($H_0: b = 3$). Coefficients were tested with a two-tailed Student's t-test. The significance of the regression was corroborated with an F-test.

Other allometric growth parameters were obtained from peer-reviewed literature. Parameters were obtained forNorth Carolina [@barbour_2011], Northern [@fogg_2013] and Southern Gulf of Mexico [@aguilarperera_2016], the Southern Mexican Caribbean [@sabidoitza_2016], Little Cayman [@edwards_2014], Jamaica [@chin_2016], Bonaire [@deleon_2013] and Costa Rica [@sandel_2015]. Additionally, parameters from Fishbase [@froese_website_2016] were also included in this study. When available, information on sampling methods, gender differentiation, location, and depth ranges of each study was retrieved. Whenever gender was not specified, it was assumed that the results were presented for both genders pooled together. During the review process, it was noticed that some papers indistinctly use $a$ to report either the ponderal index in eq. 1 or the y-intercept ($c$) in eq. 3, which might sometimes be overlooked. Furthermore, some studies report their parameters as mm-to-gr conversions, but a rapid evaluation of such parameters indicates that they were estimated as cm-to-gr conversions. Here, all parameters are reported as TL(mm) to TW(gr) conversions. When required, values from other studies are transformed for consistency.

Since standard errors or confidence intervals were not provided for many studies, it was impossible to test for statistical differences across studies. Thus expected biomass for each organism was calculated with the growth parameters estimated in this study, as well as those retrieved from additional literature. All expected biomass values were divided by the observed biomass to obtain a ratio that allowed rapid identification of over- or underestimation with respect to the observed biomass. Therefore, values lower than 1 indicate that the parameters underestimate biomass, and values larger than 1 indicate overestimation. Median and mean expected to observed biomass ratios were calculated for each study.

All hypothesis testing was performed with a confidence level of $\alpha = 0.01$ in R version 3.4.0 [@rcore_2017]. Data wrangling was performed with the tidyverse package [@tidyverse_2017]. Maps were created with a mix of functions from the sp [@sp_2017], rgdal [@rgdal_2017], tmap [@tmap_2017], and tmaptools [@tmaptools_2017] packages. Calculation of heteroskedastic-robust standard errors was done with the sandwich [@sandwich_2014] and lmtest [@lmtest_2002] packages. Models were manipulated with with the broom package [@broom_2017]. Stargazer [@stargazer_2015], knitr [@knitr_2017], and kableExtra [@kableExtra_2017] were used to produce the tables, and RefManageR [@RefManager_2014] was used to keep track of citations. The manuscript was written in r markdown [@rmarkdown_2017] and processed with the knitr package [@knitr_2017]. Raw data and code used in this work is available at [github.com/jcvdav/lionfish_biometry](github.com/jcvdav/lionfish_biometry).

@@

Organism TL ranged between 34 and 310 mm and TW between 0.3 and 397.7 gr. The smallest organism (TL = `r formatC(min(fish$Length), digits = 2, format = "f")` mm) was also the lightest organism (TW = `r formatC(min(fish$Weight), digits = 2, format = "f")` gr). However, the largest organism (TL = `r formatC(max(fish$Length), digits = 2, format = "f")` mm) was not the heaviest (TW = `r formatC(fish$Weight[fish$Length == max(fish$Length)], digits = 2, format = "f")` gr), and the heaviest organism (TW = `r formatC(max(fish$Weight), digits = 2, format = "f")` gr) was `r formatC(fish$Length[fish$Weight == max(fish$Weight)], digits = 2, format = "f")` mm in total length. Kernell density plots (Fig. 2) show the distribution for TL and TW. Both measures were positively skewed, with skewness of `r formatC(skewness(fish$Length), digits = 2, format = "f")` for TL and `r formatC(skewness(fish$Weight), digits = 2, format = "f")` for TW.

@@

wald_results <- waldtest(model, vcov = vcovHC(model, type = "HC1"))

f_string <- c("F Statistic", paste0(formatC(wald_results$F[2], digits = 2, format = "f"), "*** (df = 1; ", dim(fish)[1]-2,")"))

@@

The model adjusted to eq. 3 estimated the coefficient values at $b = `r coef(model)[[2]]`$ and $c = `r coef(model)[[1]]`$. Thus, TW (gr) can be calculated from TL (mm) as a linear equation: $log_{10}(TW) = `r coef(model)[[2]]`\times log_{10}(TL) `r coef(model)[[1]]`$ or its exponential form: $TW = `r 10^coef(model)[[1]]`\times TL^{`r coef(model)[[2]]`}$. The intercept ($c$) and slope $(b)$ were significantly different from zero ($t(107) = `r formatC(coef(model)[[1]]/robust_se[[1]], digits = 2, format = "f")`; p<0.01$ and $t(107) = `r formatC(coef(model)[[2]]/robust_se[[2]], digits = 2, format = "f")`; p<0.01$, respectively), rejecting the null hypothesis of no change. Additionally, the allometric factor ($m = b$) was significantly different from the value of isometric growth of $b = 3$ ($t(107) = `r formatC((coef(model)[[2]]-3)/robust_se[[2]], digits = 2, format = "f")`; p<0.01$), indicating that lionfish present allometric growth. More information on model fit and confidence intervals for the estimated coefficients is presented in Table 2. The relationship between Total Length and Total Weight is presented in Figure 3.

@@

```{r, include = F}
review <- read_csv("biometric_review.csv", col_types = cols(), na = "-")

review_table <- review %>% 
  select(Reference = Study, Sex = Gender, n = N, a, c, b, r2, minD, maxD, Spear) %>% 
  mutate(r2 = ifelse(is.na(r2), "-", r2),
         minD = ifelse(is.na(minD), "-", minD),
         maxD = ifelse(is.na(maxD), "-", maxD)) %>% 
  knitr::kable(row.names = F,
               digits =  c(1, 0, 6, 4, 4, 4, 0, 1, 1),
               booktabs = T,
               col.names = c("Reference", "n", "a", "c", "b", "R2", "Sex", "mDepth", "MDepth", "Spear")) %>% 
  kableExtra::kable_styling(latex_options = c("scale_down")) 
```

From the eight peer-reviewed studies including information on growth parameters for *P. volitans* and Fishbase [@froese_website_2016], `r dim(review)[1]-1` parameters were identified (Table 3). Two studies [@aguilarperera_2016; @fogg_2013] reported gender-level and pooled parameters, while the rest of the studies always presented pooled results. The smallest coefficient of determination was presented by @chin_2016 with $R^2 = `r min(review[,12], na.rm = T)`$, while @sabidoitza_2016 reported the highest value at $R^2 = `r max(review[,12], na.rm = T)`$. These studies presented information for organisms obtained at depths between `r min(review$minD, na.rm = T)` and `r max(review$maxD, na.rm = T)` m. Two studies [@aguilarperera_2016; @chin_2016] explicitly stated that their organisms were sampled with pole spears. Five studies [@sandel_2015; @barbour_2011; @fogg_2013; @edwards_2014; @sabidoitza_2016] mentioned that some of their organisms were obtained with pole spears (or other type of harpoon). A single study [@deleon_2013] did not specify how samples were obtained.

Parameters from models fit to males or females exclusively tend to have a higher steepness (*i.e.* higher allometric parameter), with mean $\pm$ standard deviation values of $b = `r formatC(mean(review$b[review$Gender == "Male"]), digits = 2, format = "f")` \pm `r formatC(sd(review$b[review$Gender == "Male"]), digits = 2, format = "f")`$ and $b = `r formatC(mean(review$b[review$Gender == "Female"]), digits = 2, format = "f")` \pm `r formatC(sd(review$b[review$Gender == "Female"]), digits = 2, format = "f")`$ males and females respectively, compared to parameters from models for pooled genders with a mean $\pm$ standard deviation value of $b = `r formatC(mean(review$b[review$Gender == "Both"]), digits = 2, format = "f")` \pm `r formatC(sd(review$b[review$Gender == "Both"]), digits = 2, format = "f")`$. In the case of the ponderal index ($a$) and its $log_{10}$ transformed parameter ($c$), values were higher for parameters for pooled genders. Figure 4 shows the predicted weights for organisms within the size range of these study using the 14 parameters previously described.

@@

# Discussion
